---
title: "Codes for Figure 4"
author: "Xiang Zhang"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(broom)
library(ggplot2)
library(forcats)
library(rstan)
library(tidybayes)
source("./R/customized_functions.R")
source("./R/OCRbayes.R")
```

# Load OCR-stats OCR data

```{r}
yepez.ocr <- read.delim("Yepez_sample_ocr_data.txt")
yepez.meta <- read.delim("Yepez_sample_meta_data.txt")

yepez.df <- yepez.ocr %>% 
  as_tibble()
```

## Plates containing cell line 61818

```{r}
plates_contain_61818 <- find_host_plate(yepez.df, target_cell_line = "61818")

plates_contain_61818
```

There are `r length(plates_contain_76065)` plates that contained cell line 76065. 

## Missing data

```{r}
yepez.df %>% 
  dplyr::filter(plate_id %in% plates_contain_61818, 
                is.na(OCR))
```

```{r include=FALSE}
plate_to_remove <- yepez.df %>% 
  dplyr::filter(plate_id %in% plates_contain_61818, is.na(OCR)) %>% 
  pull(plate_id)
```


Here, we can see that there is a missing OCR value for well H5 of plate `r plate_to_remove`. 
Since dealing missing observation is not the focus of this study, I will remove all the data related this well from the analysis. 

```{r}
dat61818 <- yepez.df %>% 
  dplyr::filter(plate_id %in% plates_contain_61818, 
                Cell_line %in% c("NHDF", "61818"), 
                !((plate_id == "20130930_1109") & (well == "G3")))
```

## How many wells (per group) are there in each plate

```{r}
dat61818 %>% 
  dplyr::select(plate_id, well, Cell_line) %>% 
  distinct() %>% 
  group_by(plate_id, Cell_line) %>% 
  summarize(number_of_wells = n()) %>% 
  knitr::kable()
```

# First experiment 

```{r}
dat.lev0 <- dat61818 %>% 
  dplyr::filter(plate_id %in% c("20130930_1109"))

experiment_setup <- tag_experimental_setup(dat.lev0)

dat_list_for_stan <- collect_data_for_stan(dat.lev0)

scenario_1_plate <- sampling(my_stan_mod, data = dat_list_for_stan, chains = 4L, cores = 4L, control = list(max_treedepth = 10))

respiration_1_plate <- convert_ocr_to_respiration(scenario_1_plate, experiment_setup)
#respiration_1_plate$scenario <- "1 plate"
respiration_1_plate$plate_id <- "20130930_1109"

respiration_1_plate %>% 
  spread(Cell_line, val) %>% 
  dplyr::mutate(diff_in_respir = log2(`61818`) - log2(NHDF)) %>% 
  dplyr::group_by(what) %>% 
  summarize(post_mean = mean(diff_in_respir), 
            low_mean = quantile(diff_in_respir, probs = 0.025), 
            high_mean = quantile(diff_in_respir, probs = 0.975))
```

# Repeat the experiment

```{r}
dat.lev0 <- dat61818 %>% 
  dplyr::filter(plate_id %in% c("20140202_1849"))

experiment_setup <- tag_experimental_setup(dat.lev0)

dat_list_for_stan <- collect_data_for_stan(dat.lev0)

scenario_2_plates <- sampling(my_stan_mod, data = dat_list_for_stan, chains = 4L, cores = 4L, control = list(max_treedepth = 10))

respiration_2_plates <- convert_ocr_to_respiration(scenario_2_plates, experiment_setup)
respiration_2_plates$plate_id <- "20140202_1849"

respiration_2_plates %>% 
  spread(Cell_line, val) %>% 
  dplyr::mutate(diff_in_respir = log2(`61818`) - log2(NHDF)) %>% 
  dplyr::group_by(what) %>% 
  summarize(post_mean = mean(diff_in_respir), 
            low_mean = quantile(diff_in_respir, probs = 0.025), 
            high_mean = quantile(diff_in_respir, probs = 0.975))
```

# Repeat the experiment third time

```{r}
dat.lev0 <- dat61818 %>% 
  dplyr::filter(plate_id %in% c("20140205_1620"))

experiment_setup <- tag_experimental_setup(dat.lev0)

dat_list_for_stan <- collect_data_for_stan(dat.lev0)

scenario_3_plates <- sampling(my_stan_mod, data = dat_list_for_stan, chains = 4L, cores = 4L, control = list(max_treedepth = 10, adapt_delta = 0.95))

respiration_3_plates <- convert_ocr_to_respiration(scenario_3_plates, experiment_setup)
respiration_3_plates$plate_id <- "20140205_1620"

respiration_3_plates %>% 
  spread(Cell_line, val) %>% 
  dplyr::mutate(diff_in_respir = log2(`61818`) - log2(NHDF)) %>% 
  dplyr::group_by(what) %>% 
  summarize(post_mean = mean(diff_in_respir), 
            low_mean = quantile(diff_in_respir, probs = 0.025), 
            high_mean = quantile(diff_in_respir, probs = 0.975))
```

```{r}
dat.lev0 <- dat61818 %>% 
  dplyr::filter(plate_id %in% c("20130930_1109", "20140202_1849", "20140205_1620"))

experiment_setup <- tag_experimental_setup(dat.lev0)

dat_list_for_stan <- collect_data_for_stan(dat.lev0)

scenario_3_plates <- sampling(my_stan_mod, data = dat_list_for_stan, chains = 4L, cores = 4L, control = list(max_treedepth = 10, adapt_delta = 0.95))

respiration_3_plates_combine <- convert_ocr_to_respiration(scenario_3_plates, experiment_setup)
respiration_3_plates_combine$plate_id <- "3 Plates Combined"

respiration_3_plates_combine %>% 
  spread(Cell_line, val) %>% 
  dplyr::mutate(diff_in_respir = log2(`61818`) - log2(NHDF)) %>% 
  dplyr::group_by(what) %>% 
  summarize(post_mean = mean(diff_in_respir), 
            low_mean = quantile(diff_in_respir, probs = 0.025), 
            high_mean = quantile(diff_in_respir, probs = 0.975))
```

```{r}
post_respiration <- respiration_1_plate %>% 
  bind_rows(respiration_2_plates) %>% 
  bind_rows(respiration_3_plates) %>% 
  #bind_rows(respiration_3_plates_combine) %>% 
  dplyr::mutate(plate_id = factor(plate_id))

dat61818 %>% 
  dplyr::filter(plate_id == "20130930_1109", 
                Cell_line == "NHDF", 
                Interval == "Int4") %>% 
  arrange(desc(OCR))

fig1 <- ggplot(dat61818, aes(as.factor(time), OCR)) + 
  geom_line(aes(group = well, color = factor(Cell_line, levels = c("NHDF", "61818"), 
                                             labels = c("Control", "Patient")))) + 
  scale_color_brewer("Cell line", palette = "Set2") + 
  facet_wrap(~plate_id, ncol = 1) + 
  labs(x = "Measurement cycle", y = "OCR (pmol/min)", 
       title = "A") + 
  theme_bw() + 
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 14), 
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12), 
        panel.grid = element_blank())

fig2 <- post_respiration %>% 
  dplyr::filter(what == "Maximal respiration") %>% 
  mutate(Cell_line = factor(Cell_line, labels = c("Control", "Patient"))) %>%  
  ggplot(aes(val)) + 
  geom_density(aes(color = Cell_line), size = 1) + 
  scale_color_brewer("Cell line", palette = "Set2") + 
  facet_wrap(~plate_id, ncol = 1, scales = "free_y") + 
  labs(x = "Maximal respiration (pmol/min)", 
       title = "B") + 
  theme_bw() + 
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 14), 
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12), 
        panel.grid = element_blank())

fig3 <- respiration_3_plates_combine %>% 
  dplyr::filter(what == "Maximal respiration") %>% 
  mutate(Cell_line = factor(Cell_line, labels = c("Control", "Patient"))) %>%  
  ggplot(aes(val)) + 
  geom_density(aes(color = Cell_line), size = 1) + 
  scale_color_brewer("Cell line", palette = "Set2") + 
  facet_wrap(~plate_id, ncol = 1, scales = "free_y") + 
  labs(x = "Maximal respiration (pmol/min)", 
       title = "C") + 
  theme_bw() + 
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 14), 
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12), 
        panel.grid = element_blank())

gridExtra::grid.arrange(fig1, fig2, fig3, nrow = 1)
```
